---
title: "SPICE_CoralBleaching"
output: html_document
---

###Packages
*Load  Packages
```{r setup, include=FALSE}
library(readxl)
library(lubridate)# for working with dates
library(ggplot2)  # for creating graphs
library(scales)   # to access breaks/formatting functions
library(gridExtra) # for arranging plots
library(janitor)
library(pivotaltrackR)
library(pivottabler)
library(here)
library(tidyverse)
library(plotly)
library(reshape2)
library(dplyr)
library(patchwork)
```

###Data
*Reading in the data file
```{r}
Global_Bleaching <- read.csv("Global_Coral_Bleaching - Florida.csv")
```

###Cleaning Up the Data
*Combining columns to have one column that shows MDY (no individual columns)
```{r}
Global_Bleaching <- Global_Bleaching %>%
  unite("SampleDate", Date_Day, Date_Month, Date_Year, sep = "-")
```

*Recognizing 'SampleDate' within the function dmy() 
```{r}
Global_Bleaching$SampleDate <- dmy(Global_Bleaching$SampleDate)
class(Global_Bleaching$SampleDate)
```





###Plot Percent Bleaching 
*Scatterplot for Date vs Percent Bleaching 
```{r}
Bleachplot <- plot(Global_Bleaching$SampleDate, Global_Bleaching$Percent_Bleaching,
     main = "Bleaching Percentage over Time",
     xlab = "Sample Date",
     ylab = "Bleaching Percentage (%)",
     pch = 16,  # Change shape of points
     col = "skyblue"  # Change color of points
)
```



*Change character to numeric
```{r}
class(Global_Bleaching$Percent_Bleaching)
```

```{r}
Global_Bleaching <- Global_Bleaching %>%
  mutate(Percent_Bleaching = as.numeric(Percent_Bleaching))
```

```{r}
ggplot(Global_Bleaching, aes(x=SampleDate, y=Percent_Bleaching)) + 
  geom_point()+
  labs(title="Percent Bleaching vs Time",
       x="Sample Date", y = "Percent Bleaching (%)")+
  theme_classic()  
```







###Plot Temperature 
*Scatterplot for Date vs Temperature
```{r}
Tempplot <- plot(Global_Bleaching$SampleDate, Global_Bleaching$Temperature_Celsius,
     main = "Temperature over Time",
     xlab = "Sample Date",
     ylab = "Temperature (C)",
     pch = 16,  # Change shape of points
     col = "skyblue",  # Change color of points
     ylim = c(18,34)
)
```

*Check numeric
```{r}
class(Global_Bleaching$Temperature_Celsius)
```

```{r}
ggplot(Global_Bleaching, aes(x=SampleDate, y=Temperature_Celsius)) + 
  geom_point()+
  labs(title="Temperature vs Time",
       x="Sample Date", y = "Temperature Celsius (C)")
  theme_classic()
```
```{r}
ggplot(Global_Bleaching, aes(x = SampleDate, y = Temperature_Celsius)) + 
  geom_point(color = "blue", size = 1.5) +  # Customize point color and size
  labs(title = "Temperature vs Time",
       x = "Sample Date", y = "Temperature Celsius (C)") +
  ylim(18, 34) +  # Set y-axis limits
  theme_classic()
```







#Make interactive
* EXAMPLE FROM CONNER
```{r}
interactive_bleached_plot <- ggplot(data = Global_Bleaching, aes(x = SampleDate,
                                                                 y = Percent_Bleaching,
                                                                 color = Temperature_Celsius)) +
                                                                 geom_point() +
  theme_bw()

ggplotly(interactive_bleached_plot) 
```

*MINE
```{r}
PvTplot <- ggplot(Global_Bleaching, aes(x=SampleDate, y=Percent_Bleaching, color = Temperature_Celsius)) + 
  geom_point()+
  labs(title="Percent Bleaching vs Time",
       x="Sample Date", y = "Percent Bleaching (%)")+
  theme_classic()  

ggplotly(PvTplot) 
```




