---
title: "SPICE_MarineFisheries"
output: html_document
---

###Packages
*Load Necessary Packages
```{r setup, include=FALSE}
library(readxl)
library(lubridate)# for working with dates
library(ggplot2)  # for creating graphs
library(scales)   # to access breaks/formatting functions
library(gridExtra) # for arranging plots
library(janitor)
library(pivotaltrackR)
library(pivottabler)
library(here)
library(tidyverse)
library(plotly)
library(reshape2)
library(gridExtra)
```



###Reef Fish
##RedSnapper
*Read in Data
```{r}
RedSnapper <- read.csv("SPICE_MarineFisheries_Reef - RedSnapper.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
par(mgp = c(3, 0.3, -1))
RSplot <- barplot(RedSnapper$TotalCatch,
        names.arg = RedSnapper$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "Red Snapper Total Catch",
        ylim = c(0, 9500000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```

##FloridaPompano
*Read in Data
```{r}
FloridaPompano <- read.csv("SPICE_MarineFisheries_Reef - FloridaPompano.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
par(mgp = c(3, 0.3, -1))
FPplot <- barplot(FloridaPompano$TotalCatch,
        names.arg = FloridaPompano$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "Florida Pompano Total Catch",
        ylim = c(0, 7500000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```

##GrayTriggerfish
*Read in Data
```{r}
GrayTriggerfish <- read.csv("SPICE_MarineFisheries_Reef - GrayTriggerfish.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
par(mgp = c(3, 0.3, -1))
GTplot <- barplot(GrayTriggerfish$TotalCatch,
        names.arg = GrayTriggerfish$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "Gray Triggerfish Total Catch",
        ylim = c(0, 5000000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```



##Plotting all Reef Species on one Bar Graph
```{r}
Year=c(1987, 1988, 1989, 1990, 1991, 1992, 1993, 1994, 1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018)
```

#Put data into proper format
*RedSnapper
```{r}
rstc = RedSnapper["TotalCatch"]
rstype = RedSnapper["Name"]
```

```{r}
RF1 = data.frame(rstc, rstype, Year)
```

*FloridaPompano
```{r}
fptc = FloridaPompano["TotalCatch"]
fptype = FloridaPompano["Name"]
```

```{r}
RF2 = data.frame(fptc, fptype, Year)
```

*GrayTriggerfish
```{r}
gttc=GrayTriggerfish["TotalCatch"]
gttype = GrayTriggerfish["Name"]
```

```{r}
RF3 = data.frame(gttc, gttype, Year)
```

#Combining Data Sets
```{r}
RF4=merge(RF1, RF2, by=c("TotalCatch", "Name", "Year"), all.x=T, all.y=T)
```

```{r}
RF5=merge(RF4, RF3, by=c("TotalCatch", "Name", "Year"), all.x=T, all.y=T)
```

#Plot Three Together
```{r}
ggplot(RF5, aes(x=Year, y=TotalCatch, fill=Name)) +
geom_bar(stat='identity', position='dodge') +
  labs(title = "Reef Fish Population per Year")
```



###Pelagic Fish
##GreatBarracuda
*Read in Data
```{r}
GreatBarracuda <- read.csv("SPICE_MarineFisheries_Large - GreatBarracuda.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
#par(mgp = c(4, 0.3, -1))
GBplot <- barplot(GreatBarracuda$TotalCatch,
        names.arg = GreatBarracuda$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "GreatBarracuda Total Catch",
        ylim = c(0, 3000000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```

##BlackfinTuna
*Read in Data
```{r}
BlackfinTuna <- read.csv("SPICE_MarineFisheries_Large - BlackfinTuna.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
BTplot <- barplot(BlackfinTuna$TotalCatch,
        names.arg = BlackfinTuna$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "Blackfin Tuna Total Catch",
        ylim = c(0, 500000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```

##GreaterAmberjack
*Read in Data
```{r}
GreaterAmberjack <- read.csv("SPICE_MarineFisheries_Large - GreaterAmberjack.csv")
```

*Plot a Bar graph for Year and Total Catch
```{r}
GAplot <- barplot(GreaterAmberjack$TotalCatch,
        names.arg = GreaterAmberjack$Year,
        main = "Comparison",
        xlab = "Year",
        ylab = "Greater Amberjack Total Catch",
        ylim = c(0, 2000000),
        col = "skyblue",  # Change color of bars if desired
        border = "black",  # Add black borders around bars for clarity
        las = 2  # Rotate x-axis labels vertically for better readability
)
```




##Plotting all Pelagic Species on one Bar Graph
*'Year' is already set to equal the years, dont need to repeat

#Put data into proper format
*GreatBarracuda
```{r}
gbtc = GreatBarracuda["TotalCatch"]
gbtype = GreatBarracuda["Name"]
```

```{r}
PF1 = data.frame(gbtc, gbtype, Year)
```

*BlackfinTuna
```{r}
bttc = BlackfinTuna["TotalCatch"]
bttype = BlackfinTuna["Name"]
```

```{r}
PF2 = data.frame(bttc, bttype, Year)
```

*GreaterAmberjack
```{r}
gatc=GreaterAmberjack["TotalCatch"]
gatype = GreaterAmberjack["Name"]
```

```{r}
PF3 = data.frame(gatc, gatype, Year)
```

#Combining Data Sets
```{r}
PF4 = merge(PF1, PF2, by=c("TotalCatch", "Name", "Year"), all.x=T, all.y=T)
```

```{r}
PF5 = merge(PF4, PF3, by=c("TotalCatch", "Name", "Year"), all.x=T, all.y=T)
```

#Plot Three Together
```{r}
ggplot(PF5, aes(x=Year, y=TotalCatch, fill=Name)) +
geom_bar(stat='identity', position='dodge') +
  labs(title = "Pelagic Fish Population per Year")
```
m

